---
title: "Part1: Web scraping using API"
author: "Bohan Yin"
date: "11/24/2019"
output:
  pdf_document: 
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(dplyr)
library(geonames)
library(here)
library(gapminder)
library(countrycode)
library(magrittr)
```


```{r API key}
# To run this code in your computer, please register on geonames.org, enable the web service and run this code here: options(geonamesUsername = "your username")
```


```{r readin data, include =FALSE}
countryInfo <- GNcountryInfo()
countryInfo %>%
  as_tibble()
countryData <- gapminder
```

```{r data modification, include = FALSE}
countryInfo %<>% 
  # unify the country names and continent names
  mutate(country = countrycode(countryCode, 'iso2c', 'country.name')) %>%
  mutate(continent = countrycode(country, 'country.name', 'continent'))
countryInfo
```

```{r merge data, include = FALSE}
country_merged <- left_join(countryData, countryInfo, by = c('country','continent')) # I merge both country and continent to reduce unnecessary suffix columns
```

## 1. Plot a single graph showing the overall trend  

From Figure 1, we can see that the relationship between population density and life expectancy is overall positive. As population density increases, life expectancy increases.  

```{r plot one graph,  message = FALSE, warning = FALSE}
country_merged %<>%
  # convert chr column to dbl column
  mutate(areaInSqKm = as.numeric(areaInSqKm)) %>%
  # calculate population density. We use population data from gapminder because it has different population data of countries on different years, which is more precise for this plot 
  mutate(pop_density = (pop/areaInSqKm))
country_merged %>%
  ggplot(aes(pop_density, lifeExp)) +
  geom_point(alpha = 0.2) +
  geom_smooth() +
  labs(title = 'Figure 1: Relationship between population density and life expectancy',
       x = 'Population density (per square kilometers)',
       y = 'Life expectancy') +
  # scale x axis with log to make the plot more interpretable
  scale_x_continuous(trans = 'log2')
```

\pagebreak  

## 2. Plot a single graph showing the trends by continents  

From Figure 2, we can see that both Asia and Africa show an overall positive relationship betwee population density and life expectancy. Especially Asia presents a more robust positive relationship.  

```{r one plot by continents, message = FALSE, warning = FALSE}
country_merged %>%
  ggplot(aes(pop_density, lifeExp, color = continent)) +
  geom_point(alpha = 0.2) +
  geom_smooth() +
  labs(title = 'Figure 2: Relationship between population density and life expectancy',
       x = 'Population density (per square kilometers)',
       y = 'Life expectancy') +
  # scale x axis with log to make the plot more interpretable
  scale_x_continuous(trans = 'log2') +
  scale_color_brewer(palette = 'RdYlBu')
```
\pagebreak  

## 3. Plot multiple graphs showing the trends in different continents with different x scales

From Figure 3 we can see, only Africa and Asia show a positive relationship between population density and life expectancy. Notice that Africa experiences a negative relationship between the two before population density reaches 16/srkm, and after that point it starts to show a positive relationship between population density and life expectancy. Other continents like Americas, Europe, and Oceania do not have an obvious positive relationship between population density and life expectancy.  


```{r multiple plots, message = FALSE, warning = FALSE}
country_merged %>%
  group_by(continent) %>%
  ggplot(aes(pop_density, lifeExp)) +
  geom_point(alpha = 0.2) +
  geom_smooth() +
  labs(title = 'Figure 3: Relationship between population density and life expectancy 
       (by continents)',
       x = 'Population density (per square kilometers)',
       y = 'Life expectancy') +
  scale_x_continuous(trans = 'log2') +
  # multiple plots across continents, allowing different x axis scales
  facet_wrap (~continent, scales = "free_x")
  
```

```{r include = FALSE}
devtools::session_info()
``` 

